Where Do Events Come From?

### 2018-07-16 paper.tex
...

@@

%% Title information
\title[Short Title]{Full Title}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Subtitle}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

@@

Text of paper \ldots

\cite{ceu.tecs17}

@@

%% Bibliography
\bibliography{my}

>>>>

\usepackage{xspace}
\newcommand{\CEU}{\textsc{C\'{e}u}\xspace}

@@

%% Title information
\title{Where Do Events Come From?}
\subtitle{Reactive Programming From The Ground Up}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command

\begin{abstract}
In reactive and event-based systems, execution is guided by an external
environment which produces inputs and consumes output events to and from the
applications.
Reactive languages provide dedicated syntax and semantics to deal with events
and greatly simplify the programming experience in this domain.
%
Nevertheless, the environment is typically prefabricated in a host language and
the very central concept of events is implemented externally to the reactive
language.
This separation exposes the environment as a rigid system component that
evolves in separate from the application.
%
In this work, we propose an interrupt handler primitive for a reactive language
targeting embedded systems in order to take control of the whole event loop,
from the hardware input source and back to the hardware output sink.
%
We propose the new primitive in the context of the structured reactive language
\CEU and discuss how it synergizes with the structured paradigm to provide
automatic standby for applications and hot swapping for drivers during runtime.
\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{interrupt service routine,
          sleep mode,
          synchronous reactive programming}

\maketitle

@@

Reactive applications interact continuously and in real time with the external
world through input and output devices (e.g., buttons, displays, timers, etc.).
%
These interactions are typically represented as input events flowing from the
devices to the application and as output events flowing from the application to
the devices.
%
As illustrated in Figure~\ref{fig.loop}, devices can be encapsulated as a
single component, \emph{the environment}, which controls the system execution
in an event loop:
the application sits idle waiting for an input;
the environment awakes the application on the occurrence of an input;
the application reacts to the input and generates back one or more outputs;
the application becomes idle and the loop restarts.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{loop}
\caption{ Event loop in reactive systems.
          The environment controls the application through input \& output
          events.
\label{fig.loop}
}
\end{figure}

The environment is typically implemented in a host language (e.g., C) and
controls the main event loop, invoking entry points in the reactive language
runtime on the occurrence of inputs and also receiving output calls, both
through a documented API.
%
As examples, Esterel~\cite{esterel.ieee91} relies on C for passing events
between the environment and the running program~\cite{esterel.book.compiling},
while Elm~\cite{frp.elm} uses the concept of \emph{ports}, which allows sending
out values to JavaScript as commands and listening for values as
subscriptions~\cite{frp.elm.ports}.

The event-based separation between the application and the environment is
arguably inevitable and happens at the right level of abstraction, since it
connects the application with the operating system resources through a
non-invasive API.
%
However, it involves two languages and the programmer may have to unnecessarily
deal with multiple syntaxes, incompatible type systems, and different address
spaces.
%
Furthermore, in the context of embedded systems, a proper host OS may even be
absent or lacking enough drivers, which requires more low-level intervention
from the application.

In this work, we propose
an interrupt handler primitive for a reactive language in
order to take control of the whole event loop, from the
hardware input source and back to the hardware output
sink. We propose the new primitive in the context of
the structured reactive language \CEU and discuss how
it synergizes with the structured paradigm to provide
automatic standby for applications and hot swapping
for drivers during runtime.

with different syntaxes and semantics.
For instance, , type systems, and

abstraction
    - rigid
    - two languages
        - syntax and semantics
            - integrates with features such as lexical scope, type system
    - platforms w/o OS

\newpage

\begin{figure}
\centering
\includegraphics[width=\linewidth]{sync-async}
\caption{ Event loop in reactive systems.
          The environment controls the application through input \& output
          events.
\label{fig.loop}
}
\end{figure}

what would be

output (int,int) O;

becomes

output (int x, int y) O do
end

\begin{comment}
% Esterel
https://books.google.com.br/books?id=O5zi14i0KP4C&pg=PA31&lpg=PA31&dq=esterel+external+signal+api&source=bl&ots=43EBAqIkSJ&sig=npYylZteOmZAqv4f7SXAtJO9xr8&hl=en&sa=X&ved=0ahUKEwjFlf7-3p_cAhWPrVkKHXsFDWgQ6AEIUjAG#v=onepage&q=esterel%20external%20signal%20api&f=false
Esterel relies on general-purpose languages such as C or C++ in two ways.       
First, for portability, most Esterel compilers generate C instead of            
assembly or some other executable representation.                               
Such generated code provides an interface for passing events between the        
environment and the running program.                                            
Second, Esterel allows the use of data types and functions defined externally   
(e.g., in C) to be used within an Esterel program.

% Elm
https://guide.elm-lang.org/interop/javascript.html
Okay, in Elm, any communication with JavaScript goes through a port. Think of
it like a hole in the side of your Elm program where you can send values in and
out. These work exactly like the commands and subscriptions from the
Architecture section. Sending values out to JS is a command. Listening for
values coming in from JS is a subscription. Pretty neat!
\end{comment}

- excel define formulas for input but not the input itself

output -> input

the environment sits
    inverse
    reacts to output and generates inputs

remain idle and react to the occurrence of events

@@

%% Bibliography
\bibliography{my,other}
